---
title: "tutoriel_R"
output: github_document
---
exo 1
```{r}
a <- c("lannister", "targaryen", "baratheon", "starck", "greyjoy")
length(a)
a[1:3]
b <- c("lannister", "starck")
a[-1]
sort(a)

```
exo 2
```{r}
a <- 1:100
a
a <- c(a, 200, 201, 202)
a
b <- seq(2,100, by=2)
b
```
3.3.2.1 data frame example
```{r}
data(iris)
View(iris)
str(iris)
```
what can we notice? 

3.3.2.2 build your own data frame
```{r}
date <- c("1_monday", "2_Tuesday", "3_Wednesday", "4_Thursday", "5_Friday", "6_Saturday", "7_Sunday")
is.character(date)

#temperature in deg Celcius
temperature <- c(24, 27, 25, 22, 30, 21, 28)
is.numeric(temperature)

#rain in mm
rain <- c(1, 0, 0, 5, 2, 0, 0)
is.numeric(rain)

#make data.frame
df <- data.frame(date, temperature, rain)
str(df)

#to select a column or vector
df$temperature
df[, 2]

# here we use list() instead of c() because there is pultiple class in inside row
day <- list("8_monday", 29, 1)
new_def <- rbind(df, day) #add row to a data frame
new_def

```
4 Functions
head() # to know first line of your data frame
class()# return the class of the object. Ex : data.frame, matrix, list ....
str()# return the structure of the object. Ex : numeric, factor, character....
names()# to get or set the names of an object
sum() # for addition
min() # return the minimum of the vector
max() # return the minimum of the vector
row.names() # attribute names for lines of the data frame
colnames() # attribute names for column of the data frame
apply() # Returns a vector or array or list of values obtained by applying a function to margins of an array or matrix.
```{r}
head(df)
class(df)
str(df)
apply(df[, 2:3], 2, mean)
```
exo 3 
```{r}
data("iris")

```

